### 2021-03-05 paper2/1-introduction.tex
Fix all kinds of small problems
Use include instead of import
This has the advantage that I can compile one chapter at a time.

% The problem of fair representations can also be viewed through a similar lens.
% A good toy model for this setup is the Coloured MNIST (cMNIST) dataset.
% In this example, the colour is a spurious variable which is very closely correlated with the prediction target in the training set but not in the deployment setting.
% An interpretable invariant representation in this case is the images uniform in colour. More details regarding the setup and synthesis of the dataset can be found in Section~\ref{ssec:cmnist}.

@@

\begin{figure*}[tb]
  \centering
  \subfloat[Original images.]{%
      \scalebox{0.3}{\includegraphics[width=\textwidth]{./Images/celeba/train_original_x_2.png}}%
      \label{fig:cflow_celeba_original_x}
  }
  \hfill
  \subfloat[$\bm{x}_u$ null-samples from the cFlow model.]{%
      \scalebox{0.3}{\includegraphics[width=\textwidth]{./Images/celeba/train_reconstruction_y_2.png}}%
      \label{fig:cflow_celeba_recon_y}
  }
  \hfill
  \subfloat[$\bm{x}_b$ null-samples from the cFlow model.]{%
      \scalebox{0.3}{\includegraphics[width=\textwidth]{./Images/celeba/train_reconstruction_s_2.png}}%
      \label{fig:cflow_celeba_recon_s}
  }
  \caption{
    CelebA null-samples learned by our cFlow model, with gender as the sensitive attribute.
    (a) The original, untransformed samples from the CelebA dataset
    (b) Reconstructions using only information unrelated to $s$.
    (c) Reconstruction using only information related to $\neg s$.
    The model learns to disentangle gender from the non-gender related information.
    % Attributes such as \emph{makeup} and \emph{hair length} are also often modified in the process due to inherent correlations between them and the sensitive attribute, which the intepretability of our representations allows us to easily identify.
    Note that some attributes like skin tone seem to change along with gender due to the correlation between the attributes.
    This is especially visible in images (1,1) and (3,2). Only because our representations are produced in the data-domain can we easily spot such instances of entanglement.
  }%
  \label{fig:celeba_cflow}
\end{figure*}

>>>>

% The problem of fair representations can also be viewed through a similar lens.
% A good toy model for this setup is the Coloured MNIST (cMNIST) dataset.
% In this example, the colour is a spurious variable which is very closely correlated with the prediction target in the training set but not in the deployment setting.
% An interpretable invariant representation in this case is the images uniform in colour. More details regarding the setup and synthesis of the dataset can be found in section~\ref{ssec:cmnist}.

@@

\begin{figure*}[tb]
  \centering
  \subfloat[Original images.]{%
      \scalebox{0.3}{\includegraphics[width=\textwidth]{paper2/Images/celeba/train_original_x_2.png}}%
      \label{fig:cflow_celeba_original_x}
  }
  \hfill
  \subfloat[$\bm{x}_u$ null-samples from the cFlow model.]{%
      \scalebox{0.3}{\includegraphics[width=\textwidth]{paper2/Images/celeba/train_reconstruction_y_2.png}}%
      \label{fig:cflow_celeba_recon_y}
  }
  \hfill
  \subfloat[$\bm{x}_b$ null-samples from the cFlow model.]{%
      \scalebox{0.3}{\includegraphics[width=\textwidth]{paper2/Images/celeba/train_reconstruction_s_2.png}}%
      \label{fig:cflow_celeba_recon_s}
  }
  \caption{
    CelebA null-samples learned by our cFlow model, with gender as the sensitive attribute.
    (a) The original, untransformed samples from the CelebA dataset
    (b) Reconstructions using only information unrelated to $s$.
    (c) Reconstruction using only information related to $\neg s$.
    The model learns to disentangle gender from the non-gender related information.
    % Attributes such as \emph{makeup} and \emph{hair length} are also often modified in the process due to inherent correlations between them and the sensitive attribute, which the intepretability of our representations allows us to easily identify.
    Note that some attributes like skin tone seem to change along with gender due to the correlation between the attributes.
    This is especially visible in images (1,1) and (3,2). Only because our representations are produced in the data-domain can we easily spot such instances of entanglement.
  }%
  \label{fig:celeba_cflow}
\end{figure*}

### 2021-03-05 paper2/2-related-work.tex
Fix headings and remove duplicate bib entries
Use include instead of import
This has the advantage that I can compile one chapter at a time.

\begin{figure*}[!htb]
    \centering
    \subfloat[Samples from the cMNIST training set, $\sigma=0$.]{%
        \scalebox{0.3}{\includegraphics[width=\textwidth]{./Images/cmnist/cflow_original_task_x_scale_0.png}}%
        \label{fig:cflow_cmnist_task_train}
    }
    % \hfill
    % \subfloat[Samples from the cMNIST test set with $\sigma=0.02$.]{%
    %     \scalebox{0.225}{\includegraphics[width=\textwidth]{./Images/cmnist/cflow_pretrain_x.png}}%
    %     \label{fig:cflow_cmnist_pretrain}
    % }
    \hfill
    \subfloat[$x_u$ null-samples from the cFlow model.]{%
        \scalebox{0.3}{\includegraphics[width=\textwidth]{./Images/cmnist/cflow_task_xy_scale_0.png}}%
        \label{fig:cflow_cmnist_y}
    }
    \hfill
    \subfloat[$x_b$ null-samples from the cFlow model.]{%
        \scalebox{0.3}{\includegraphics[width=\textwidth]{./Images/cmnist/cflow_task_xs_scale_0.png}}%
        \label{fig:cflow_cmnist_s}
    }
    \caption{
        Sample images from the coloured MNIST dataset problem with $10$ predefined mean colours.
        (a): Images from the spuriously correlated subpopulation where colour is a reliable signal of the digit class-label.
        (b-c): Results of running our approach realised with cFlow on the cMNIST dataset.
        The model learns to retain the shape of the digit shape while removing the relationship with colour.
        A downstream classifier is now less prone to exploiting correlations between colour and the digit label class.
    }\label{fig:cmnist}
\end{figure*}

@@

\subsection{Learning fair representations.}
%As we have alluded to, the goal of producing invariant representations is similar to that of producing \emph{fair} representations.
% In fairness problems, there is usually a \emph{sensitive attribute}, $s$ (for example, gender or race),
% that should not be used to make decisions.
Given a sensitive attribute $s$ (for example, gender or race) and inputs $\bm{x}$,
a fair representation $\bm{z}$ of $\bm{x}$ is then one for which $\bm{z} \perp s$ holds,
while ideally also being predictive of the class label $y$.
\citet{ZemWuSwePitetal13} was the first to propose the learning of fair representations which allow for transfer to new classification tasks.
More recent methods are often based on variational autoencoders (VAEs)~\citep{kingma2013auto,louizos2016variational,edwards2016censoring,beutel2017data}.
The achieved fairness of the representation can be measured with various fairness metrics.
These measure, however, usually how fair the predictions of a classifier are
and not how fair a representation is.
% defined with respect to predictions and not representations.

>>>>

\begin{figure*}[!htb]
    \centering
    \subfloat[Samples from the cMNIST training set, $\sigma=0$.]{%
        \scalebox{0.3}{\includegraphics[width=\textwidth]{paper2/Images/cmnist/cflow_original_task_x_scale_0.png}}%
        \label{fig:cflow_cmnist_task_train}
    }
    % \hfill
    % \subfloat[Samples from the cMNIST test set with $\sigma=0.02$.]{%
    %     \scalebox{0.225}{\includegraphics[width=\textwidth]{./Images/cmnist/cflow_pretrain_x.png}}%
    %     \label{fig:cflow_cmnist_pretrain}
    % }
    \hfill
    \subfloat[$x_u$ null-samples from the cFlow model.]{%
        \scalebox{0.3}{\includegraphics[width=\textwidth]{paper2/Images/cmnist/cflow_task_xy_scale_0.png}}%
        \label{fig:cflow_cmnist_y}
    }
    \hfill
    \subfloat[$x_b$ null-samples from the cFlow model.]{%
        \scalebox{0.3}{\includegraphics[width=\textwidth]{paper2/Images/cmnist/cflow_task_xs_scale_0.png}}%
        \label{fig:cflow_cmnist_s}
    }
    \caption{
        Sample images from the coloured MNIST dataset problem with $10$ predefined mean colours.
        (a): Images from the spuriously correlated subpopulation where colour is a reliable signal of the digit class-label.
        (b-c): Results of running our approach realised with cFlow on the cMNIST dataset.
        The model learns to retain the shape of the digit shape while removing the relationship with colour.
        A downstream classifier is now less prone to exploiting correlations between colour and the digit label class.
    }\label{fig:cmnist}
\end{figure*}

@@

\subsection{Learning fair representations.}
%As we have alluded to, the goal of producing invariant representations is similar to that of producing \emph{fair} representations.
% In fairness problems, there is usually a \emph{sensitive attribute}, $s$ (for example, gender or race),
% that should not be used to make decisions.
Given a sensitive attribute $s$ (for example, gender or race) and inputs $\bm{x}$,
a fair representation $\bm{z}$ of $\bm{x}$ is then one for which $\bm{z} \perp s$ holds,
while ideally also being predictive of the class label $y$.
\citet{zemel2013learning} was the first to propose the learning of fair representations which allow for transfer to new classification tasks.
More recent methods are often based on variational autoencoders (VAEs)~\citep{kingma2013auto,louizos2016variational,edwards2016censoring,beutel2017data}.
The achieved fairness of the representation can be measured with various fairness metrics.
These measure, however, usually how fair the predictions of a classifier are
and not how fair a representation is.
% defined with respect to predictions and not representations.

### 2021-03-05 paper2/3-method.tex
Fix all kinds of small problems
Use include instead of import
This has the advantage that I can compile one chapter at a time.

\section{Interpretable Invariances by Null-Sampling}\label{ssec:general}
% \begin{figure}
%     \centering
%     \includegraphics[width=0.4\textwidth]{./Figures/diagram.pdf}
%     \caption{Training procedure using the cFlow model for illustrative purposes.}%
%     \label{fig:training_diagram}
% \end{figure}
\begin{figure*}[tb]
    \centering
    \hfill
    \subfloat[cFlow model.]{%
        \scalebox{0.33}{\includegraphics[width=\textwidth]{./Figures/inn_diagram_u.pdf}}%
        \label{fig:inn_diagram}
    }
    \hfill
    \subfloat[cVAE model.]{%
        \scalebox{0.4}{\includegraphics[width=\textwidth]{./Figures/cvae_diagram_u.pdf}}%
        \label{fig:cvae_diagram}
    }
    \hfill
    \caption{
        Training procedure for our models. $x$: input, $s$: sensitive attribute, $z_u$: de-biased representation, $x_u$: de-biased version of the input in the data domain.
        The red bar indicates a gradient reversal layer, and $\stackrel{\rightarrow}{0}$ the null-sampling operation.
    }%
    \label{fig:model-diagrams}
\end{figure*}

@@

We now summarise the training procedure; an outline for the invertible network model (cFlow) can be seen in Fig.~\ref{fig:inn_diagram}.
First, the encoder network $f$ is trained on ($X^\mathit{rep}, S^\mathit{rep}$), during the first phase.
The trained network is then used to encode the training set,
taking in $\bm{x}$ and producing the representation, $\bm{z}_u$, decorrelated from the spurious variable.
The encoded dataset can then be used to train any off-the-shelf classifier safely, with information about the spurious variable having been absorbed by some auxiliary component $\mathcal{B}$.
In the case of the conditional VAE (cVAE) model,
$\mathcal{B}$ takes the form of the decoder subnetwork, which reconstructs the data conditional on a one-hot encoding of $s$,
while for the invertible network $\mathcal{B}$ is realised as a partition of the feature map $\bm{z}$
(such that $\bm{z} = [\bm{z}_u, \bm{z}_b]$), given the bijective constraint.
Thus, the classifier cannot take the shortcut of learning $s$ and instead must learn how to predict $y$ directly.
Obtaining the $s$-invariant representations, $\bm{x}_u$, in the data domain
is simply a matter of replacing the $\mathcal{B}$ component of the decoder's input for the cVAE,
and $\bm{z}_b$ for cFlow, with a zero vector of equivalent size.
We refer to this procedure used to generate $\bm{x}_u$ as \emph{null-sampling} (here, with respect to $\bm{z}_b)$.

>>>>

\section{Interpretable Invariances by Null-Sampling}\label{ssec:general}
% \begin{figure}
%     \centering
%     \includegraphics[width=0.4\textwidth]{paper2/Figures/diagram.pdf}
%     \caption{Training procedure using the cFlow model for illustrative purposes.}%
%     \label{fig:training_diagram}
% \end{figure}
\begin{figure*}[tb]
    \centering
    \hfill
    \subfloat[cFlow model.]{%
        \scalebox{0.33}{\includegraphics[width=\textwidth]{paper2/Figures/inn_diagram_u.pdf}}%
        \label{fig:inn_diagram}
    }
    \hfill
    \subfloat[cVAE model.]{%
        \scalebox{0.4}{\includegraphics[width=\textwidth]{paper2/Figures/cvae_diagram_u.pdf}}%
        \label{fig:cvae_diagram}
    }
    \hfill
    \caption{
        Training procedure for our models. $x$: input, $s$: sensitive attribute, $z_u$: de-biased representation, $x_u$: de-biased version of the input in the data domain.
        The red bar indicates a gradient reversal layer, and $\stackrel{\rightarrow}{0}$ the null-sampling operation.
    }%
    \label{fig:model-diagrams}
\end{figure*}

@@

We now summarise the training procedure; an outline for the invertible network model (cFlow) can be seen in fig.~\ref{fig:inn_diagram}.
First, the encoder network $f$ is trained on ($X^\mathit{rep}, S^\mathit{rep}$), during the first phase.
The trained network is then used to encode the training set,
taking in $\bm{x}$ and producing the representation, $\bm{z}_u$, decorrelated from the spurious variable.
The encoded dataset can then be used to train any off-the-shelf classifier safely, with information about the spurious variable having been absorbed by some auxiliary component $\mathcal{B}$.
In the case of the conditional VAE (cVAE) model,
$\mathcal{B}$ takes the form of the decoder subnetwork, which reconstructs the data conditional on a one-hot encoding of $s$,
while for the invertible network $\mathcal{B}$ is realised as a partition of the feature map $\bm{z}$
(such that $\bm{z} = [\bm{z}_u, \bm{z}_b]$), given the bijective constraint.
Thus, the classifier cannot take the shortcut of learning $s$ and instead must learn how to predict $y$ directly.
Obtaining the $s$-invariant representations, $\bm{x}_u$, in the data domain
is simply a matter of replacing the $\mathcal{B}$ component of the decoder's input for the cVAE,
and $\bm{z}_b$ for cFlow, with a zero vector of equivalent size.
We refer to this procedure used to generate $\bm{x}_u$ as \emph{null-sampling} (here, with respect to $\bm{z}_b)$.

### 2021-03-05 paper2/4-experiments.tex
Use include instead of import
This has the advantage that I can compile one chapter at a time.

\section{Experiments}
\noindent We present experiments to demonstrate that the null-sampled representations are in fact invariant to $s$
while still allowing a classifier to predict $y$ from them.
We run our cVAE and cFlow models on the coloured MNIST (cMNIST) and CelebA dataset,
which we artificially bias, first describing the sampling procedure we follow to do so for non-synthetic datasets.
As baselines we have the model of~\citet{kim2019learning} (Ln2L) and the same CNN used to evaluate the cFlow and cVAE models
but with the unmodified images as input (CNN).
For the cFlow model we adopt a Glow-like architecture~\citep{KinDha18},
while both subnetworks of the cVAE model comprise gated convolutions~\citep{van2016conditional}, where the encoding size is $256$.
For cMNIST, we construct the Ln2L baseline according to its original description, for CelebA,
we treat it as an augmentation of the baseline CNN's objective function.
%More
Detailed information regarding model architectures can be found in Appendix~\ref{sec:architectures} and~\ref{sec:optimisation-details}.%
\footnote{The code can be found at \url{https://github.com/predictive-analytics-lab/nifr}.}
% \begin{figure*}[!htb]
%   \centering
%   %
%   \subfloat[CelebA]{\scalebox{0.495}{\includegraphics[width=\textwidth]{./results/celeba/celeba_acc_DP.pdf}}
%   \label{fig:celeba_chart}
%   }%
%   %\hfill
%   \subfloat[UCI Adult]{\scalebox{0.495}{\includegraphics[width=\textwidth]{./Figures/nosinn_celeba_multiplot_all_landscape_Smiling.pdf}}
%   \label{fig:adult_chart}
%   }%
%   \caption{
%     Performance of our approach on the CelebA and UCI Adult datasets.
%     (Left of each) We show the accuracy of a downstream classifier trained on representations learned by our model.
%     (Right of each) We show the Demographic Parity of a model trained on our representations (the lower the better).
%   }\label{fig:chart}
% \end{figure*}
\begin{figure}[tb]
    \centering
    \includegraphics[width=0.7\textwidth]{./Figures/nosinn_celeba.pdf}
    \caption{
        Performance of our model for different targets (mixing factor $\eta=0$).
        Left: \emph{Smiling} as target, right: \emph{high cheekbones}.
        \emph{DP diff} measures fairness with respect to demographic parity.
        A perfectly fair model has a \emph{DP diff} of 0.
    }%
    \label{fig:celeba-targets}
\end{figure}

\begin{figure}[tb]
    \centering
    \includegraphics[width=0.85\textwidth]{./Figures/nosinn_celeba_multiplot_all_landscape_Smiling.pdf}
    \caption{
        Performance of our model for the target ``smiling'' for different mixing factors $\eta$.
        \emph{DP diff} measures fairness with respect to demographic parity.
        A perfectly fair model has a \emph{DP diff} of 0, thus the closer to top-left the better it is in terms of we accuracy-fairness trade-off.
        Only values $\eta=0$ and $\eta=1$ correspond to the scenario of a strongly biased training set.
        The results for $0.1\leq \eta\leq 0.9$ are to confirm that our model does not harm performance for non-biased training sets.
    }%
    \label{fig:celeba-multiplot}
\end{figure}

@@

\begin{figure}[tb]
    \centering
    % \includegraphics[width=0.9\textwidth]{./results/cmnist/cmnist_results.pdf}
    % \includegraphics[width=0.9\textwidth]{./Figures/cmnist_new.pdf}
    \includegraphics[width=0.7\textwidth]{./Figures/cmnist_new_no_hgr.pdf}
    \caption{
        % Accuracy and correlation (HGR) of predictions $\hat{y}$ and $s$ of our approach in comparison with other baseline models on the cMNIST dataset, for different standard deviations ($\sigma$) for the colour sampling.
        Accuracy of our approach in comparison with other baseline models on the cMNIST dataset, for different standard deviations ($\sigma$) for the colour sampling.
    }%
    \label{fig:cmnist_chart}
\end{figure}
\subsection{Experimental results}
% \noindent Our model can be applied to tabular data and image data alike.
% Due to space constraints, the experiments with the tabular dataset \emph{UCI Adult} are situated in the Appendix.
%To illustrate the effectiveness and generality of our framework, we apply our models to a tabular dataset, in the UCI Adult dataset,
We report the results from two image datasets.
cMNIST, a synthetic dataset, is a good starting point for evaluating our model due to the direct control we have over the biasing.
CelebA, on the other hand, is a more practical and challenging example.
% To illustrate the generality of our approach,
We also test our method on a tabular dataset, the Adult dataset.
%For full details regarding the architectures of our models and their optimisation, see the Appendix.

\paragraph{cMNIST.}
The coloured MNIST (cMNIST) dataset is a variant of the MNIST dataset in which the digits are coloured.
In the training set, the colours have a one-to-one correspondence with the digit class.
In the test set (and the representative set), colours are assigned randomly.
The colours are drawn from Gaussians with 10 different means.
We follow the colourisation procedure outlined by~\citet{kim2019learning}, with the mean colour values selected so as to be maximally dispersed.
The full list of such values can be found in Appendix~\ref{sec:color-details}.
We produce multiple variants of the cMNIST dataset corresponding to different standard deviations $\sigma$ for the colour sampling:
$\sigma \in \{0.00, 0.01, ..., 0.05 \}$.
% The images are symmetrically zero-padded to a size of $32 \times 32$ to allow for an additional downsampling stage in the cFlow model.
% No additional data augmentation is used.
% 24,000 of the 36,000 training samples were set aside for representative, the remaining 10,000 for training the downstream classifier.
% We continue to use 10,000 samples usually reserved for testing on vanilla MNIST for that same purpose.

@@

\paragraph{CelebA.}
To evaluate the effectiveness of our framework on real-world image data we use the CelebA dataset~\citep{liu2015faceattributes}, consisting of 202,599 celebrity images.
These images are annotated with various binary physical  attributes, including ``gender'', ``hair color'', ``young'', etc, from which we  select our sensitive and target attributes.
The images are centre cropped and resized to $64\times64$, as is standard practice.
For our experiments, we designate ``gender'' as the sensitive attribute,
and ``smiling'' and ``high cheekbones'' as target attributes.
We chose gender as the sensitive attribute as it a common sensitive attribute in the fairness literature.
For the target attributes, we chose attributes that are harder to learn than gender and which do not correlate too strongly with gender in the dataset
(``wearing lipstick'' for example being an attribute too closely correlated with gender).
% The sampling and evaluation procedure is identical to that conducted for the UCI Adult dataset.
% We evaluate the performance of all models for mixing factors ($\eta$) of value $\{0, 0.1, ..., 1\}$. 
The model is trained on the representative set (normal subset of CelebA)
and is then used to encode the artificially biased training set and the test set.
The results for the most strongly biased training set ($\eta=0$) can be found in fig.~\ref{fig:celeba-targets}.
Our method outperforms the baselines in accuracy and fairness.

@@

To show that our method can handle multinomial, as well as binary, sensitive attributes, we also conduct experiments with $s=\textrm{hair color}$ as a ternary attribute (``Blonde'', ``Black'', ``Brown''), excluding ``Red'' because of the paucity of samples and the noisiness of their labels. The results for these experiments can be found in section~\ref{sec:additional-results}.
% Looking at the produced invariant representations,
% The failure of the cVAE to model complex interactions between $x_u$ and $s$ is brought to bear here.
% Due to the balancing between the adversarial and reconstruction losses,
% information relevant to the downstream task of predicting ``smiling'' is lost during the pre-training phase,
% leading to poor performance at all values of $\eta$ except the most extreme ones,
% for which it does reasonably well at mitigating the sampling bias (Fig.~\ref{fig:celeba_chart}).
% Meanwhile, cFlow does not suffer this drawback -- the images remain sharp
% (see Fig.~\ref{fig:cflow_celeba_recon_y} in comparison to Fig. \ref{fig:cvae_celeba_recon_y})
% and the accuracy obtained by training on them is comparable to the baseline at intermediary values -- and at the same time matches and outperforms the cVAE at $\eta = 0$ and $\eta = 1$, respectively.
% For these same values,
% incorporating the mutual information loss proposed by~\cite{kim2019learning} into the classifier's loss function greatly harmed performance measured by accuracy;
% the DP is zero due to all predictions being negative.

% \begin{wrapfigure}{l}{0.6\linewidth}
\begin{figure}[tb]
  \centering
  \includegraphics[width=0.6\textwidth]{./Figures/nosinn_adult_multiplot_mini_diff.pdf}
  \caption{
      Results for the \textsc{Adult} dataset.
      The $x$-axis corresponds to the difference in positive rates.
      An ideal result would occupy the \textsc{top-left}.
  }%
  \label{fig:adult-chart}
\end{figure}
% \end{wrapfigure}

@@

% \begin{figure}
%   \centering
%   \includegraphics[width=0.7\textwidth]{./Figures/nosinn_adult_multiplot_mini_diff.pdf}
%   \caption{
    %   Results for the \textsc{Adult} dataset.
    %   The $x$-axis corresponds to the difference in positive rates.
    %   An ideal result would occupy the \textsc{top-left}.
%   }%
%   \label{fig:adult-chart}
% \end{figure}

>>>>

\section{Experiments}
\noindent We present experiments to demonstrate that the null-sampled representations are in fact invariant to $s$
while still allowing a classifier to predict $y$ from them.
We run our cVAE and cFlow models on the coloured MNIST (cMNIST) and CelebA dataset,
which we artificially bias, first describing the sampling procedure we follow to do so for non-synthetic datasets.
As baselines we have the model of~\citet{kim2019learning} (Ln2L) and the same CNN used to evaluate the cFlow and cVAE models
but with the unmodified images as input (CNN).
For the cFlow model we adopt a Glow-like architecture~\citep{KinDha18},
while both subnetworks of the cVAE model comprise gated convolutions~\citep{van2016conditional}, where the encoding size is $256$.
For cMNIST, we construct the Ln2L baseline according to its original description, for CelebA,
we treat it as an augmentation of the baseline CNN's objective function.
%More
Detailed information regarding model architectures can be found in Appendix~\ref{sec:architectures} and~\ref{sec:optimisation-details}.%
\footnote{The code can be found at \url{https://github.com/predictive-analytics-lab/nifr}.}
% \begin{figure*}[!htb]
%   \centering
%   %
%   \subfloat[CelebA]{\scalebox{0.495}{\includegraphics[width=\textwidth]{./results/celeba/celeba_acc_DP.pdf}}
%   \label{fig:celeba_chart}
%   }%
%   %\hfill
%   \subfloat[UCI Adult]{\scalebox{0.495}{\includegraphics[width=\textwidth]{paper2/Figures/nosinn_celeba_multiplot_all_landscape_Smiling.pdf}}
%   \label{fig:adult_chart}
%   }%
%   \caption{
%     Performance of our approach on the CelebA and UCI Adult datasets.
%     (Left of each) We show the accuracy of a downstream classifier trained on representations learned by our model.
%     (Right of each) We show the Demographic Parity of a model trained on our representations (the lower the better).
%   }\label{fig:chart}
% \end{figure*}
\begin{figure}[tb]
    \centering
    \includegraphics[width=0.7\textwidth]{paper2/Figures/nosinn_celeba.pdf}
    \caption{
        Performance of our model for different targets (mixing factor $\eta=0$).
        Left: \emph{Smiling} as target, right: \emph{high cheekbones}.
        \emph{DP diff} measures fairness with respect to demographic parity.
        A perfectly fair model has a \emph{DP diff} of 0.
    }%
    \label{fig:celeba-targets}
\end{figure}

\begin{figure}[tb]
    \centering
    \includegraphics[width=0.85\textwidth]{paper2/Figures/nosinn_celeba_multiplot_all_landscape_Smiling.pdf}
    \caption{
        Performance of our model for the target ``smiling'' for different mixing factors $\eta$.
        \emph{DP diff} measures fairness with respect to demographic parity.
        A perfectly fair model has a \emph{DP diff} of 0, thus the closer to top-left the better it is in terms of we accuracy-fairness trade-off.
        Only values $\eta=0$ and $\eta=1$ correspond to the scenario of a strongly biased training set.
        The results for $0.1\leq \eta\leq 0.9$ are to confirm that our model does not harm performance for non-biased training sets.
    }%
    \label{fig:celeba-multiplot}
\end{figure}

@@

\begin{figure}[tb]
    \centering
    % \includegraphics[width=0.9\textwidth]{./results/cmnist/cmnist_results.pdf}
    % \includegraphics[width=0.9\textwidth]{paper2/Figures/cmnist_new.pdf}
    \includegraphics[width=0.7\textwidth]{paper2/Figures/cmnist_new_no_hgr.pdf}
    \caption{
        % Accuracy and correlation (HGR) of predictions $\hat{y}$ and $s$ of our approach in comparison with other baseline models on the cMNIST dataset, for different standard deviations ($\sigma$) for the colour sampling.
        Accuracy of our approach in comparison with other baseline models on the cMNIST dataset, for different standard deviations ($\sigma$) for the colour sampling.
    }%
    \label{fig:cmnist_chart}
\end{figure}
\subsection{Experimental results}
% \noindent Our model can be applied to tabular data and image data alike.
% Due to space constraints, the experiments with the tabular dataset \emph{UCI Adult} are situated in the Appendix.
%To illustrate the effectiveness and generality of our framework, we apply our models to a tabular dataset, in the UCI Adult dataset,
We report the results from two image datasets.
cMNIST, a synthetic dataset, is a good starting point for evaluating our model due to the direct control we have over the biasing.
CelebA, on the other hand, is a more practical and challenging example.
% To illustrate the generality of our approach,
We also test our method on a tabular dataset, the Adult dataset.
%For full details regarding the architectures of our models and their optimisation, see the Appendix.

\paragraph{cMNIST.}
The coloured MNIST (cMNIST) dataset is a variant of the MNIST dataset in which the digits are coloured.
In the training set, the colours have a one-to-one correspondence with the digit class.
In the test set (and the representative set), colours are assigned randomly.
The colours are drawn from Gaussians with 10 different means.
We follow the colourisation procedure outlined by~\citet{kim2019learning}, with the mean colour values selected so as to be maximally dispersed.
The full list of such values can be found in section~\ref{sec:color-details}.
We produce multiple variants of the cMNIST dataset corresponding to different standard deviations $\sigma$ for the colour sampling:
$\sigma \in \{0.00, 0.01, ..., 0.05 \}$.
% The images are symmetrically zero-padded to a size of $32 \times 32$ to allow for an additional downsampling stage in the cFlow model.
% No additional data augmentation is used.
% 24,000 of the 36,000 training samples were set aside for representative, the remaining 10,000 for training the downstream classifier.
% We continue to use 10,000 samples usually reserved for testing on vanilla MNIST for that same purpose.

@@

\paragraph{CelebA.}
To evaluate the effectiveness of our framework on real-world image data we use the CelebA dataset~\citep{liu2015faceattributes}, consisting of 202,599 celebrity images.
These images are annotated with various binary physical  attributes, including ``gender'', ``hair colour'', ``young'', etc, from which we  select our sensitive and target attributes.
The images are centre cropped and resized to $64\times64$, as is standard practice.
For our experiments, we designate ``gender'' as the sensitive attribute,
and ``smiling'' and ``high cheekbones'' as target attributes.
We chose gender as the sensitive attribute as it a common sensitive attribute in the fairness literature.
For the target attributes, we chose attributes that are harder to learn than gender and which do not correlate too strongly with gender in the dataset
(``wearing lipstick'' for example being an attribute too closely correlated with gender).
% The sampling and evaluation procedure is identical to that conducted for the UCI Adult dataset.
% We evaluate the performance of all models for mixing factors ($\eta$) of value $\{0, 0.1, ..., 1\}$. 
The model is trained on the representative set (normal subset of CelebA)
and is then used to encode the artificially biased training set and the test set.
The results for the most strongly biased training set ($\eta=0$) can be found in fig.~\ref{fig:celeba-targets}.
Our method outperforms the baselines in accuracy and fairness.

@@

To show that our method can handle multinomial, as well as binary, sensitive attributes, we also conduct experiments with $s=\textrm{hair colour}$ as a ternary attribute (``Blonde'', ``Black'', ``Brown''), excluding ``Red'' because of the paucity of samples and the noisiness of their labels. The results for these experiments can be found in section~\ref{sec:additional-results}.
% Looking at the produced invariant representations,
% The failure of the cVAE to model complex interactions between $x_u$ and $s$ is brought to bear here.
% Due to the balancing between the adversarial and reconstruction losses,
% information relevant to the downstream task of predicting ``smiling'' is lost during the pre-training phase,
% leading to poor performance at all values of $\eta$ except the most extreme ones,
% for which it does reasonably well at mitigating the sampling bias (Fig.~\ref{fig:celeba_chart}).
% Meanwhile, cFlow does not suffer this drawback -- the images remain sharp
% (see Fig.~\ref{fig:cflow_celeba_recon_y} in comparison to Fig. \ref{fig:cvae_celeba_recon_y})
% and the accuracy obtained by training on them is comparable to the baseline at intermediary values -- and at the same time matches and outperforms the cVAE at $\eta = 0$ and $\eta = 1$, respectively.
% For these same values,
% incorporating the mutual information loss proposed by~\cite{kim2019learning} into the classifier's loss function greatly harmed performance measured by accuracy;
% the DP is zero due to all predictions being negative.

% \begin{wrapfigure}{l}{0.6\linewidth}
\begin{figure}[tb]
  \centering
  \includegraphics[width=0.6\textwidth]{paper2/Figures/nosinn_adult_multiplot_mini_diff.pdf}
  \caption{
      Results for the \textsc{Adult} dataset.
      The $x$-axis corresponds to the difference in positive rates.
      An ideal result would occupy the \textsc{top-left}.
  }%
  \label{fig:adult-chart}
\end{figure}
% \end{wrapfigure}

@@

% \begin{figure}
%   \centering
%   \includegraphics[width=0.7\textwidth]{paper2/Figures/nosinn_adult_multiplot_mini_diff.pdf}
%   \caption{
    %   Results for the \textsc{Adult} dataset.
    %   The $x$-axis corresponds to the difference in positive rates.
    %   An ideal result would occupy the \textsc{top-left}.
%   }%
%   \label{fig:adult-chart}
% \end{figure}
